Openness and computational reproducibility in plant pathology: where do we stand and a way forward

### 2022-10-13 analysis/paper/paper.Rmd
replace *e.g.* with e.g. throughout

Open science and reproducibility is becoming more mainstream with many funding agencies expecting data to be available upon the conclusion of the research project [*e.g.*, @government_of_canada_2016; @vannoorden2017; @ARC2018; @EUCommission2022; @GatesFoundation2022], journals in the field promoting the sharing of data [@DelPonte2020], and other scientists becoming interested in sharing their raw data and even lab notebook contents [@Wald2010] or drawing attention to the lack of code and data that make the published work less useful [@Barton2022].

@@

In the biological sciences it is not always possible to use identical test material or perhaps the time or resources are not available for full reproducibility, *e.g.*, field trials that span years and locations or complex glasshouse experiments.
Therefore, we will follow Peng's [-@Peng2009] definition that provides clear guidelines for a minimum standard of "reproducible research":

@@

To understand where we stand as a discipline regarding open science and reproducible research, we surveyed a broad selection of articles to represent a wide swath of publications to evaluate our collective status.
We hand-picked 21 journals which we felt represented research publications in the field of plant pathology (Table \@ref(tab:journals)) that encompassed a range of subject matter foci, applied and fundamental work, country of origin and ranking metrics, *e.g.*, quartile range or citation index, to represent the discipline.
The aim was to gather as complete an overview of the status of computational reproducibility in plant pathology journals as possible and avoid bias in the findings by skewing towards high-impact journals that may have a greater influence.
From those 21 journals we randomly selected 450 articles published from 2012 to 2021.
Using R [@RCT2022], two lists were created, the first was a list of the 21 journals and the second was a list of the evaluators that were evaluating the articles for reproducibility.
Initially there were four evaluators, later a fifth was added and the list was recreated.
There were three evaluation events that took place to increase the yearly coverage of the evaluations.
During each scoring event, each evaluator was randomly assigned 50 unique articles to evaluate during an evaluation event.

@@

Articles were then manually selected by visiting the journals' websites and selecting the articles within which the randomly assigned page numbers fell, *i.e.*, if the page number was 32, the article that started on page 28 and ended on page 35 was selected as it contained page 32.
In cases where an article was not suitable, *e.g.*, a review or otherwise not related to plant pathology or the randomly assigned numbers for that journal and year fell within the same article, the next article in that journal was selected until a suitable article was found.

@@

Each article was rated on a 0 -- 3 scale for its data and code availability.
'Code availability' rated how easily and openly available the computational methods used in the article were.
Scores were assigned as '0' was 'Not available or not mentioned in the publication'; '1' was 'Available upon request to the author; '2' was 'Online, but inconvenient or non-permanent (*e.g.*, login needed, paywall, FTP server, personal lab website that may disappear, or may have already disappeared)'; and '3' was 'Freely available online to anonymous users for foreseeable future (*e.g.*, archived using Zenodo, dataverse or university library or some other proper archiving system)'; 'NA' indicates that no code was created to conduct the work that was detectable.
Second, the 'Data availability' evaluated how freely and openly available the data presented in the article were.
This was evaluated as '0' was 'Not available or not mentioned in the publication'; '1' was 'Available upon request to the author; '2' was 'Online, but inconvenient or non-permanent (*e.g.*, login needed, paywall, FTP server, personal lab website that may disappear, or may have already disappeared)'; and '3' was 'Freely available online to anonymous users for foreseeable future (*e.g.*, archived using Zenodo, dataverse or university library or some other proper archiving system)'; 'NA' indicates that no data were generated, *e.g.*, a methods paper.

@@

In cases where partial computational materials or data were made available, the score was assigned the lowest score reflecting the completeness of the availability, all or nothing.
That is, in cases where some data such as sequences were deposited in a database, *e.g.*, GenBank, but other data that was collected and used were not made available, the score assigned was '0'.

@@

The methods for selecting articles meant that only page numbers between 1 and 150 were selected.
This approach assumed that there was no effect or bias on reproducibility based on the time of year that an article was published since most journals start with page number one at the beginning of the year and some journals, *e.g.*, *Phytopathology*® have focus issues with invited authors.
In spite of some publications being from special issues, there were no detectable differences in the reproducibility of the articles as defined and measured by this article, so the assumption appears to be correct.

@@

The first layer is including tables of raw data or code with the paper as a supplemental file or even within the paper as tables where possible.
This is suitable for studies that may have a small data set or simple analysis or for demonstrations purposes as @Madden2015 demonstrate in their discussion regarding the use of $P$ values in statistical analyzes where they supply an *e-**X**tra* with reproducible examples for readers to refer to.
@Hill2019 also shared code and example data to reproduce the results of the paper as supplementary materials, citing all packages used and the versions.
The scripts provided used an R package, *checkpoint* [@Ooi2022], to provide a mechanism to help ensure reproducibility by installing the package versions used by the authors that were necessary to reproduce the work rather than defaulting to the latest versions, which may cause failures due to changes in the package code-bases.
However, the scripts did not run unhindered as the supplementary materials suggested without changes to the data.
But with some modifications to the data, and file name changes, the scripts ran allowing examples of the research to be reproduced using the definition of @Peng2009.
One drawback to this approach is that journals are often not equipped to handle code, *e.g.*, script files, that may be developed as a part of the research process and so they are often archived as PDF or Word documents which hinder the ability to easily ingest and start working with them or if they are provided in a native text format, they do not render via the web properly.
Though this should be possible given some effort from the publishing journal to share simple text files rather than binary formats along with the proper instructions and handling of these file formats rather than a one-size-fits all approach that we commonly see for supplementary materials.
For example, the R scripts for @Hill2019 were provided as supplementary materials but do not actually appear in the browser window when requested, which may confuse readers and makes downloading them more difficult.
While this allows the reader to quickly view the extra materials and a DOI is assigned as a part of the article itself, the data or code are not readily findable and accessible through a searchable database.
Furthermore, in many cases this does not allow prompt access to the data and running the code because of a journal paywall.

@@

The third layer is the use of proper code repositories such as a library resource, or code repository (*e.g.* GitHub or GitLab) and data repositories (*e.g.*, FigShare, Zenodo or OSF.io), allowing for the deposition and updating of code, figures, data preprints or any other materials that support the article itself while providing a DOI and citation.
As an example, Sparks et al. [-@Sparks2011; -@Sparks2014, @Sparks2016] used FigShare to provide models, data and code necessary to replicate model development and the subsequent study on the effects of climate change on potato late blight.
Similarly, @carleson2019population hosted the code for reproducing a population genomic analysis of *Phytophthora plurivora* on GitHub, while providing all data on OSF.io (the Open Science Framework).
@Lehner2017 used GitHub to host a code repository of their research compendium website with data and a reproducible report that explains in detail all steps of the analysis and the R code for conducting a meta-analysis for assessing heterogeneity in the relationship between white mold incidence and soybean yield and between incidence and soybean yield (<https://emdelponte.github.io/paper-white-mold-meta-analysis/>).
The website clearly demonstrates the analysis to readers and uses R so that anyone can easily replicate the work.
Using a public code repository resource allows other researchers to easily contact the authors by opening "Issues" and report bugs or ask questions in an open forum that are not as straightforward when the data are provided as supplementary material.

Taking this approach even further, packaging the full analysis in a containerized software application, *e.g.*, Docker is a way to help ensure computational reproducibility but at the expense of added complexity.
Docker is an open-source containerization platform that enables users to package several applications and an operating system into containers, thereby standardizing the executable components by combining application source code (or analyzes) with the operating system required to run that analysis on a user's computer.
However, there are drawbacks to using Docker.
It can be difficult to understand for a new user and new platforms, like the Apple M1 chip, may not be fully supported, which hinder the ability to share the container.
But, in most cases using an open-source language like R, Julia or Python allows us to share our work in a fashion where we know that the analysis will run exactly the same on every computer.
For more on using Docker for reproducible research, we would refer readers to @Nuest2020.
As an example of this approach, @Khaliq2020 provided a research compendium as a Docker container with a DOI and a full R package.
This enables readers to fully replicate their analysis of *Ascochyta rabiei* conidia dispersal in chickpea using the data collected.
It also allows for more in-depth investigation by stepping through other points where weather data were investigated and various models were fit before deciding on the best fit and recreate any figures as published in the article [@Khaliq2020a].
When this layer is employed, tools such as Binder, <https://mybinder.org>, can be used that allow readers and reviewers to launch an interactive session in their web browser and interact with the data and rerun the analysis in an RStudio instance or Jupyter notebooks, a "web application for creating and sharing computational documents" [@jupyter2022], as Mioroni et al. used [-@Miorini2018; -@Miorini2019].
@Kamvar2015 took a slightly different approach by including all files necessary for the analysis and most output files in a repository [@Kamvar2014] that also included an installable R package that was used for the original analysis.
While there are many other methods, these two approaches illustrate some of the best-practices where the data and other files were deposited in repositories with DOIs and reproducibility issues were addressed by using R packages to handle dependencies and other versioning issues making the work more portable.

@@

The first layer, as reported, is an essential step that is not substituted by the other practices and eventually researchers fail to provide sufficient description or correct citations.
When making science more open and reproducible, methods, software used (this includes items such as R, Julia or Python packages that were directly used in the analysis or production of the paper, etc.) should be cited properly.
Provide a full description of all equipment used, *e.g.*, "a Spectrum Technologies Watchdog 2700 weather station was used to record wind speed, direction, rainfall, temperature and relative humidity at one-hour intervals".
This allows end-users to identify what was used and identify the methods used more accurately.
Just as importantly, this acknowledges the contributions of others whose works were instrumental in our research.
This also helps to ensure that researchers can reconstruct what they did more easily since good notes exist and documentation and be able to identify if something changes, *e.g.*, a package version, what effect it had on the research.

@@

Text files for saving small sets of data are preferable.
Data that are saved in binary formats such as PDF files are difficult to reuse because they are not easily machine readable.
In many cases, data sets are small enough and curated in spreadsheets, which should be saved as a plain text file, *e.g.*, comma separated (CSV) or tab separated (TSV) files.
This also helps to ensure that the data are reusable.
Larger data sets may warrant the use of a proper database like lightweight personal databases, *e.g.*, SQLite or DuckDB, larger more robust databases, *e.g.*, MariaDB, PostgreSQL or a specialized database like GenBank which provide users with several benefits but two important benefits to mention here are (i) data redundancy, ensuring no records are duplicated and (ii) data consistency, ensuring that all records in a data set are recorded in the same format for every observation.
Databases such a GenBank are preferable for molecular data ensuring data integrity and machine readability.
While databases may offer many advantages for speed and data integrity, the trade-off is that they are more complex to set up and administer, especially for a small data set.

>>>>

Open science and reproducibility is becoming more mainstream with many funding agencies expecting data to be available upon the conclusion of the research project [e.g., @government_of_canada_2016; @vannoorden2017; @ARC2018; @EUCommission2022; @GatesFoundation2022], journals in the field promoting the sharing of data [@DelPonte2020], and other scientists becoming interested in sharing their raw data and even lab notebook contents [@Wald2010] or drawing attention to the lack of code and data that make the published work less useful [@Barton2022].

@@

In the biological sciences it is not always possible to use identical test material or perhaps the time or resources are not available for full reproducibility, e.g., field trials that span years and locations or complex glasshouse experiments.
Therefore, we will follow Peng's [-@Peng2009] definition that provides clear guidelines for a minimum standard of "reproducible research":

@@

To understand where we stand as a discipline regarding open science and reproducible research, we surveyed a broad selection of articles to represent a wide swath of publications to evaluate our collective status.
We hand-picked 21 journals which we felt represented research publications in the field of plant pathology (Table \@ref(tab:journals)) that encompassed a range of subject matter foci, applied and fundamental work, country of origin and ranking metrics, e.g., quartile range or citation index, to represent the discipline.
The aim was to gather as complete an overview of the status of computational reproducibility in plant pathology journals as possible and avoid bias in the findings by skewing towards high-impact journals that may have a greater influence.
From those 21 journals we randomly selected 450 articles published from 2012 to 2021.
Using R [@RCT2022], two lists were created, the first was a list of the 21 journals and the second was a list of the evaluators that were evaluating the articles for reproducibility.
Initially there were four evaluators, later a fifth was added and the list was recreated.
There were three evaluation events that took place to increase the yearly coverage of the evaluations.
During each scoring event, each evaluator was randomly assigned 50 unique articles to evaluate during an evaluation event.

@@

Articles were then manually selected by visiting the journals' websites and selecting the articles within which the randomly assigned page numbers fell, *i.e.*, if the page number was 32, the article that started on page 28 and ended on page 35 was selected as it contained page 32.
In cases where an article was not suitable, e.g., a review or otherwise not related to plant pathology or the randomly assigned numbers for that journal and year fell within the same article, the next article in that journal was selected until a suitable article was found.

@@

Each article was rated on a 0 -- 3 scale for its data and code availability.
'Code availability' rated how easily and openly available the computational methods used in the article were.
Scores were assigned as '0' was 'Not available or not mentioned in the publication'; '1' was 'Available upon request to the author; '2' was 'Online, but inconvenient or non-permanent (e.g., login needed, paywall, FTP server, personal lab website that may disappear, or may have already disappeared)'; and '3' was 'Freely available online to anonymous users for foreseeable future (e.g., archived using Zenodo, dataverse or university library or some other proper archiving system)'; 'NA' indicates that no code was created to conduct the work that was detectable.
Second, the 'Data availability' evaluated how freely and openly available the data presented in the article were.
This was evaluated as '0' was 'Not available or not mentioned in the publication'; '1' was 'Available upon request to the author; '2' was 'Online, but inconvenient or non-permanent (e.g., login needed, paywall, FTP server, personal lab website that may disappear, or may have already disappeared)'; and '3' was 'Freely available online to anonymous users for foreseeable future (e.g., archived using Zenodo, dataverse or university library or some other proper archiving system)'; 'NA' indicates that no data were generated, e.g., a methods paper.

@@

In cases where partial computational materials or data were made available, the score was assigned the lowest score reflecting the completeness of the availability, all or nothing.
That is, in cases where some data such as sequences were deposited in a database, e.g., GenBank, but other data that was collected and used were not made available, the score assigned was '0'.

@@

The methods for selecting articles meant that only page numbers between 1 and 150 were selected.
This approach assumed that there was no effect or bias on reproducibility based on the time of year that an article was published since most journals start with page number one at the beginning of the year and some journals, e.g., *Phytopathology*® have focus issues with invited authors.
In spite of some publications being from special issues, there were no detectable differences in the reproducibility of the articles as defined and measured by this article, so the assumption appears to be correct.

@@

The first layer is including tables of raw data or code with the paper as a supplemental file or even within the paper as tables where possible.
This is suitable for studies that may have a small data set or simple analysis or for demonstrations purposes as @Madden2015 demonstrate in their discussion regarding the use of $P$ values in statistical analyzes where they supply an *e-**X**tra* with reproducible examples for readers to refer to.
@Hill2019 also shared code and example data to reproduce the results of the paper as supplementary materials, citing all packages used and the versions.
The scripts provided used an R package, *checkpoint* [@Ooi2022], to provide a mechanism to help ensure reproducibility by installing the package versions used by the authors that were necessary to reproduce the work rather than defaulting to the latest versions, which may cause failures due to changes in the package code-bases.
However, the scripts did not run unhindered as the supplementary materials suggested without changes to the data.
But with some modifications to the data, and file name changes, the scripts ran allowing examples of the research to be reproduced using the definition of @Peng2009.
One drawback to this approach is that journals are often not equipped to handle code, e.g., script files, that may be developed as a part of the research process and so they are often archived as PDF or Word documents which hinder the ability to easily ingest and start working with them or if they are provided in a native text format, they do not render via the web properly.
Though this should be possible given some effort from the publishing journal to share simple text files rather than binary formats along with the proper instructions and handling of these file formats rather than a one-size-fits all approach that we commonly see for supplementary materials.
For example, the R scripts for @Hill2019 were provided as supplementary materials but do not actually appear in the browser window when requested, which may confuse readers and makes downloading them more difficult.
While this allows the reader to quickly view the extra materials and a DOI is assigned as a part of the article itself, the data or code are not readily findable and accessible through a searchable database.
Furthermore, in many cases this does not allow prompt access to the data and running the code because of a journal paywall.

@@

The third layer is the use of proper code repositories such as a library resource, or code repository (e.g. GitHub or GitLab) and data repositories (e.g., FigShare, Zenodo or OSF.io), allowing for the deposition and updating of code, figures, data preprints or any other materials that support the article itself while providing a DOI and citation.
As an example, Sparks et al. [-@Sparks2011; -@Sparks2014, @Sparks2016] used FigShare to provide models, data and code necessary to replicate model development and the subsequent study on the effects of climate change on potato late blight.
Similarly, @carleson2019population hosted the code for reproducing a population genomic analysis of *Phytophthora plurivora* on GitHub, while providing all data on OSF.io (the Open Science Framework).
@Lehner2017 used GitHub to host a code repository of their research compendium website with data and a reproducible report that explains in detail all steps of the analysis and the R code for conducting a meta-analysis for assessing heterogeneity in the relationship between white mold incidence and soybean yield and between incidence and soybean yield (<https://emdelponte.github.io/paper-white-mold-meta-analysis/>).
The website clearly demonstrates the analysis to readers and uses R so that anyone can easily replicate the work.
Using a public code repository resource allows other researchers to easily contact the authors by opening "Issues" and report bugs or ask questions in an open forum that are not as straightforward when the data are provided as supplementary material.

Taking this approach even further, packaging the full analysis in a containerized software application, e.g., Docker is a way to help ensure computational reproducibility but at the expense of added complexity.
Docker is an open-source containerization platform that enables users to package several applications and an operating system into containers, thereby standardizing the executable components by combining application source code (or analyzes) with the operating system required to run that analysis on a user's computer.
However, there are drawbacks to using Docker.
It can be difficult to understand for a new user and new platforms, like the Apple M1 chip, may not be fully supported, which hinder the ability to share the container.
But, in most cases using an open-source language like R, Julia or Python allows us to share our work in a fashion where we know that the analysis will run exactly the same on every computer.
For more on using Docker for reproducible research, we would refer readers to @Nuest2020.
As an example of this approach, @Khaliq2020 provided a research compendium as a Docker container with a DOI and a full R package.
This enables readers to fully replicate their analysis of *Ascochyta rabiei* conidia dispersal in chickpea using the data collected.
It also allows for more in-depth investigation by stepping through other points where weather data were investigated and various models were fit before deciding on the best fit and recreate any figures as published in the article [@Khaliq2020a].
When this layer is employed, tools such as Binder, <https://mybinder.org>, can be used that allow readers and reviewers to launch an interactive session in their web browser and interact with the data and rerun the analysis in an RStudio instance or Jupyter notebooks, a "web application for creating and sharing computational documents" [@jupyter2022], as Mioroni et al. used [-@Miorini2018; -@Miorini2019].
@Kamvar2015 took a slightly different approach by including all files necessary for the analysis and most output files in a repository [@Kamvar2014] that also included an installable R package that was used for the original analysis.
While there are many other methods, these two approaches illustrate some of the best-practices where the data and other files were deposited in repositories with DOIs and reproducibility issues were addressed by using R packages to handle dependencies and other versioning issues making the work more portable.

@@

The first layer, as reported, is an essential step that is not substituted by the other practices and eventually researchers fail to provide sufficient description or correct citations.
When making science more open and reproducible, methods, software used (this includes items such as R, Julia or Python packages that were directly used in the analysis or production of the paper, etc.) should be cited properly.
Provide a full description of all equipment used, e.g., "a Spectrum Technologies Watchdog 2700 weather station was used to record wind speed, direction, rainfall, temperature and relative humidity at one-hour intervals".
This allows end-users to identify what was used and identify the methods used more accurately.
Just as importantly, this acknowledges the contributions of others whose works were instrumental in our research.
This also helps to ensure that researchers can reconstruct what they did more easily since good notes exist and documentation and be able to identify if something changes, e.g., a package version, what effect it had on the research.

@@

Text files for saving small sets of data are preferable.
Data that are saved in binary formats such as PDF files are difficult to reuse because they are not easily machine readable.
In many cases, data sets are small enough and curated in spreadsheets, which should be saved as a plain text file, e.g., comma separated (CSV) or tab separated (TSV) files.
This also helps to ensure that the data are reusable.
Larger data sets may warrant the use of a proper database like lightweight personal databases, e.g., SQLite or DuckDB, larger more robust databases, e.g., MariaDB, PostgreSQL or a specialized database like GenBank which provide users with several benefits but two important benefits to mention here are (i) data redundancy, ensuring no records are duplicated and (ii) data consistency, ensuring that all records in a data set are recorded in the same format for every observation.
Databases such a GenBank are preferable for molecular data ensuring data integrity and machine readability.
While databases may offer many advantages for speed and data integrity, the trade-off is that they are more complex to set up and administer, especially for a small data set.
